---
title: "STA22"
output: html_notebook
---

```{r}
library(VGAM)
library(ggplot2)
library(MASS)
library(tidyverse)
library(caret)
library(nnet)
library(corrplot)
library(mlogit)
library(caret)
library(SciViews)
library(graphics)
library(ROCR)
library(ggplot2)
library(ggpubr)
library(mltools)
library(psych)
```
age : age
sex : sex (1: male 0:female)
cp : chest pain -> 0 : typical angina, 1:atypical angina, 2: non-anginal pain, 3:asymptomatic (least serious)
trestbps : resting blood pressure (in mm Hg on admission to hospital)
chol : serum cholestoral [mg/dl]
fbs : fasting blood sugar>120 mg/dl (1=true 0=False)--> a fasting blood sugar leval<100 is normal, 100<120: prediabetes, >125 : diabetes
restecg : resting elctrocardiographic results--> value 0--> normal, value-->1 abnormal, value-->2 left ventricular hypertrophy
thalach: maximum heart rate achieved
exang : exercise induced angina 1:yes 0:no
oldpeak : ST depression induced by exercice relative to test--> measured in J-point , <0.5 considererd as accepted , >0.5 pathological
slope : slope of peak exercise--> 0 upslopping, 1 flat,2 downslopping
ca : number of major vessels colored by flouroscopy
thal : Thalassemia (less hemoglobin than normal)



```{r}
PATH_TO_FILE="C:\\Users\\a.lance\\Documents\\MASTER\\STA211\\STA212\\hearth_disease.csv"
```

```{r}
ds<-read.csv(PATH_TO_FILE,sep=";")
ds<-na.omit(ds)
```

#suppression du dataset
```{r}
ds<-subset(ds,ds$chol<=350)
```


# transformation des données catégoriques
```{r}
ds$cp_typical_angina<-ifelse(ds$cp==0,1,0)
ds$cp_atypical_angina<-ifelse(ds$cp==1,1,0)
ds$cp_no_ang_pain<-ifelse(ds$cp==2,1,0)
ds$cp_assymp<-ifelse(ds$cp==3,1,0)
#
# ds$slope_up<-ifelse(ds$slope==0,1,0)
# ds$slope_flat<-ifelse(ds$slope==1,1,0)
# ds$slope_down<-ifelse(ds$slope==2,1,0)
#
ds$restecg_normal<-ifelse(ds$restecg==0,1,0)
ds$restecg_abnormal<-ifelse(ds$restecg==1,1,0)
ds$restecg_ventricular_hp<-ifelse(ds$restecg==2,1,0)
#
ds$thal_normal<-ifelse(ds$thal==0,1,0)
ds$thal_fixed_def<-ifelse(ds$restecg==1,1,0)
ds$thal_rev_def<-ifelse(ds$restecg==2,1,0)

```

# transformation variable cholesterol binaire
```{r}
ds$chol_normal<-ifelse(ds$chol<190,1,0)
```


```{r}
ds<-ds[ , -which(names(ds) %in% c("cp","restecg","thal"))]
ds$target<-as.factor(ds$target)
```


## Distribution de la données

```{r}
continuous_parameters<-c("age" ,"trestbps","chol","thalach","oldpeak")
for (var in continuous_parameters){
print(ggplot(ds,aes(x=ds[[var]])) + 
    geom_histogram(aes(y=..density..), binwidth=.5,colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")
     +labs(y= "density", x = var)
    )}
```
## variables binaires

```{r}
binary_parameters<-c("sex" ,
                     "fbs",
                     "exang",
                     "cp_typical_angina",
                     "cp_atypical_angina",
                     "cp_no_ang_pain",
                     "cp_assymp",
                     "restecg_normal",
                     "restecg_abnormal",
                     "restecg_ventricular_hp",
                     "thal_normal",
                     "thal_fixed_def",
                     "thal_rev_def",
                     "chol_normal")

mcc_phi<-as.numeric()
for (var in binary_parameters){
print(ggplot(ds, aes(x=ds[[var]])) + geom_histogram(binwidth=.5)+scale_x_continuous(n.breaks = 2)+labs(y= "count", x = var))
  cont.table<-table(ds[[var]],ds$target)
  mcc_phi<-append(mcc_phi,phi(cont.table,digits = 3))
}
ds.heatmap<-data.frame(phi.coeff=mcc_phi,row.names = binary_parameters)
ds.heatmap

```


```{r}
ordinal_parameters<-c("ca")
for (var in ordinal_parameters){
print(ggplot(ds, aes(x=ds[[var]])) + geom_histogram(binwidth=.5)+scale_x_continuous(n.breaks = 2)+labs(y= "density", x = var))}
```




```{r}
continuous_parameters<-c("age" ,"trestbps","chol","thalach","oldpeak")
for (var in continuous_parameters){
    ds_<-ds[c('target',var)]
    ds_$target<-as.factor(ds_$target)
    print(
    ggdensity(ds_, x = var,
    add = "mean", 
    rug = TRUE,
    color = "target", palette = c("#00AFBB", "#E7B800"))
    )
    }
```

## Impact des variables sur target


```{r}
parameters<-colnames(ds)
for (var in parameters){
print(ggplot(ds, aes(x=target, y=ds[[var]])) + geom_boxplot()
      +labs(y= var, x = 'Maladie cardiaque diagnostiquée')
      )
}
```
#Calcul des coefficients de mathews pour les variables binaires

```{r}
ggplot(ds, aes(x = slope, y = target,)) +
  geom_col(position = "fill")
```



```{r}
set.seed(1234)
sample <- sample(c(TRUE, FALSE), nrow(ds), replace=TRUE, prob=c(0.7,0.3))
ds.train  <- ds[sample, ]
ds.test   <- ds[!sample, ]
```

# Set univariate model for each parameters
```{r}
#parameters<-c("age","sex","cp" ,"trestbps","chol","fbs","restecg","thalach","exang","oldpeak","slope" ,"ca","thal")
parameters<-colnames(ds[ , -which(names(ds) %in% c("target"))])
p_values = numeric()
p_values2 = numeric()
z_value<-numeric()
Beta<-numeric()
GList<-numeric()
wald.test.list<-numeric()
SE.list<-numeric()
LCI.list<-numeric()
UCI.list<-numeric()
for (param in parameters){
  
  fmod<-glm(reformulate(c(param), response='target'),data = ds.train,family = "binomial")
  nmod <- glm(target~1,data = ds.train, family = 'binomial')
  G= -2*ln(as.numeric(exp(logLik(nmod)))/as.numeric(exp(logLik(fmod))))
  GList<- append(GList,G)
  p_values2<-append(p_values2,pchisq(G, 1,lower.tail = FALSE))
  Beta1<-as.numeric(fmod$coefficients[2])
  Beta<-append(Beta,Beta1)
  SE<-as.numeric(sqrt(diag(vcov(fmod)))[2])
  SE.list<-append(SE.list,SE)
  z_value<-abs(Beta1/SE)
  wald_test_p_value<-2*pnorm(z_value,lower.tail = FALSE)
  wald.test.list<-append(wald.test.list,wald_test_p_value)
  LCI.list<-append( LCI.list,Beta1-1.96*SE)
  UCI.list<-append( UCI.list,Beta1+1.96*SE)
}

```

```{r}
df.results<-data.frame(row.names =parameters )
df.results$coeffs<-Beta
df.results$p_value<-p_values2
df.results$p_valueWT<-wald.test.list
df.results$SE<-SE.list
df.results$LCI<-LCI.list
df.results$UCI<-UCI.list
df.results
```

Le paramètres fbs/cp_assymp/chol_normal est exclus

```{r}
#parameters<-c("age","sex","cp" ,"trestbps","chol","restecg","thalach","exang","oldpeak","slope" ,"ca","thal")
parameters<-colnames(ds[ , -which(names(ds) %in% c("target","fbs","cp_assymp","chol_normal"))])
glm.model<-glm(reformulate(parameters, response='target'),data = ds.train,family = "binomial")
coeff.list.2<-as.numeric(glm.model$coefficients)
summary(glm.model)
```


paramètres à exclure : age/cp_atypical_angina/cp_typical_angina/cp_no_ang_pain/cp_assymp/rest_ecg_normal/restecg_abnormal/rest_ecg_ventricular_hp/thal_normal/thal_fixed_df/thal_rev_def

```{r}
parameters<-colnames(ds[ , -which(names(ds) %in% c("target",
                                                   "fbs",
                                                   "age",
                                                   "cp_typical_angina",
                                                   "cp_assymp",
                                                   "cp_no_ang_pain",
                                                   "cp_atypical_angina",
                                                   "restecg_normal",
                                                   "restecg_abnormal",
                                                   "restecg_ventricular_hp",
                                                   "thal_normal",
                                                   "thal_fixed_def",
                                                   "thal_rev_def",
                                                   "chol_normal"
                                                   ))])

glm.model<-glm(reformulate(parameters, response='target'),data = ds.train,family = "binomial")
coeff.list.3<-as.numeric(glm.model$coefficients)
summary(glm.model)
```
```{r}
parameters_excluded<-c(
                       "target",
                       "fbs",
                       "age",
                       "cp_typical_angina",
                       "cp_assymp",
                       "cp_no_ang_pain",
                       "cp_atypical_angina",
                       "restecg_normal",
                       "restecg_abnormal",
                       "restecg_ventricular_hp",
                       "thal_normal",
                       "thal_fixed_def",
                       "thal_rev_def",
                       "chol_normal"
                       )

parameters<-colnames(ds[ , -which(names(ds) %in%parameters_excluded)])

for (p in parameters_excluded){
  print(p)
  new_param<-append(parameters,p)
  glm.model2<-glm(reformulate(new_param, response='target'),data = ds.train,family = "binomial")
  print(summary(glm.model2))
}
```

# Vérification logits lineaire pour variables continues

```{r}
probabilities <- predict(glm.model, type = "response")

predictors<-c('thalach','oldpeak','trestbps',"chol")
ds.train.logit<-ds.train[predictors]
# Bind the logit and tidying the data for plot
ds.train.logit <- ds.train.logit %>%
  mutate(logit = log(probabilities/(1-probabilities))) %>%
  gather(key = "predictors", value = "predictor.value", -logit)


ggplot(ds.train.logit, aes(logit, predictor.value))+
  geom_point(size = 0.5, alpha = 0.5) +
  geom_smooth(method = "loess") + 
  theme_bw() + 
  facet_wrap(~predictors, scales = "free_y")

```
## Intéractions potentielles

```{r}
parameters<-colnames(ds[ , -which(names(ds) %in% c("target",
                                                   "fbs",
                                                   "age",
                                                   "cp_assymp",
                                                   "cp_no_ang_pain",
                                                   "cp_atypical_angina",
                                                   "slope_up",
                                                   "slope_down",
                                                   "restecg_normal",
                                                   "restecg_abnormal",
                                                   "restecg_ventricular_hp",
                                                   "thal_normal",
                                                   "thal_fixed_def",
                                                   "thal_rev_def"
                                                   ))])

p.value.list<-as.numeric()
coeff.list<-as.numeric()
combination<-as.character()
for (i in 1:length(parameters))
#for (i in parameters)
  {
    for (j in i:length(parameters)) 
        {
        if (parameters[i]!=parameters[j])
          {
            factor1= paste(parameters[i],parameters[j],sep='*')
            factor2= paste0(parameters,collapse="+")
            factor<-paste0(factor2,"+",factor1)
            print(factor)
            glm.fit <- glm(as.formula(paste0("target~", factor)), data=ds.train,family = "binomial")
            summary_<-summary(glm.fit)
            coeff.list<-append(coeff.list,summary_$coefficients[11,1])
            p.value.list<-append(p.value.list,summary_$coefficients[11,4])
            combination<-append(combination,factor1)
          }
        }
  }
  

```

```{r}
df.results<-data.frame(row.names =combination )
df.results$coeffs<-coeff.list
df.results$p_value<-p.value.list
df.results<-subset(df.results,p_value<0.01)
df.results
```
## Sélection des intéractions

```{r}
significant_interactions<-rownames(df.results)
initial_parameters<-c("sex","trestbps","chol","thalach","exang","oldpeak","ca","cp_typical_angina","slope")

all_intercation_combinations<-do.call("c",lapply(seq_along(significant_interactions),function(i) utils::combn(significant_interactions,i,FUN=list)))
for (c in all_intercation_combinations)
  {
    formula_<-paste0("target~",paste0(initial_parameters,collapse="+"),"+",paste0(c,collapse = '+'))
    glm.final_ <- glm(as.formula(formula_), data=ds.train,family = "binomial")
    print(summary(glm.final_))
  }
```
## Modèle final

Coefficients:
                   Estimate Std. Error z value Pr(>|z|)    
(Intercept)        7.273765   1.450991   5.013 5.36e-07 ***
sex               -1.822474   0.287657  -6.336 2.36e-10 ***
trestbps          -0.025326   0.006512  -3.889 0.000100 ***
chol              -0.018365   0.003455  -5.316 1.06e-07 ***
thalach            0.021193   0.006139   3.452 0.000557 ***
exang             -0.793576   0.268250  -2.958 0.003093 ** 
oldpeak           -0.435157   0.131983  -3.297 0.000977 ***
ca                -0.890254   0.141238  -6.303 2.91e-10 ***
cp_typical_angina -1.820965   0.243964  -7.464 8.39e-14 ***
slope              0.643559   0.229758   2.801 0.005094 ** 
chol_normal       -1.654223   0.539005  -3.069 0.002148 ** 
ca:chol_normal     0.679102   0.299187   2.270 0.023218 *  
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1



```{r}
final_parameters<-c('sex','trestbps','chol','thalach','exang','oldpeak','ca','cp_typical_angina','slope','chol_normal')
formula_<-paste0("target~",paste0(initial_parameters,collapse="+"),"+",paste0(c("ca","chol_normal"),collapse = '*'))
glm.final_ <- glm(as.formula(formula_), data=ds.train,family = binomial(link="logit"))
print(summary(glm.final_))
```


```{r}
confint(glm.final_)
```

## ROC CURVE

```{r}
pred<-predict(glm.final_,ds.train[parameters],type="response")
pred_ <- prediction(pred, ds.train$target)
perf <- performance(pred_,"tpr","fpr")
plot(perf,colorize=TRUE)
```

```{r}
auc_ROCR<-performance(pred_,measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
auc_ROCR
```
# Test performance

```{r}
pred<-predict(glm.final_,ds.test[parameters],type="response")
treshold<-0.5
Yfac   <- ds.test$target
yhatFac<-cut(pred,breaks = c(-Inf,treshold,Inf),lables=c('0','1'))
contingency.table<-table(Yfac,yhatFac)
addmargins(contingency.table)
sum(diag(contingency.table)) / sum(contingency.table)
```

## Interprétation
# ODD ratio

```{r}
dich.variables<-c('sex','fbs')
contin.variables<-c('chol','thalach','oldpeak','trestbps')
comb.list<-as.character()
delta.beta.list<-as.numeric()
for (d.var in dich.variables){
  for (c.var in contin.variables){
    comb.list<-append(comb.list,(paste0(c(d.var,c.var),collapse='*')))
    univar.model<-glm(as.formula(paste0('target~',d.var)),data=ds,family = binomial(link="logit"))
    multi.model<-glm(as.formula(paste0('target~',paste0(c(d.var,c.var),collapse='+'))),data=ds,family = binomial(link="logit"))
    multi.interact.model<-glm(as.formula(paste0('target~',paste0(c(d.var,c.var),collapse='+'),'+',paste0(c(d.var,c.var),collapse='*'))),data=ds,family = binomial(link="logit"))
    Beta1.univar.model<-as.numeric(univar.model$coefficients[d.var])
    Beta1.multivar.model<-as.numeric(multi.model$coefficients[d.var])
    delta.Beta.perc<-(Beta1.univar.model-Beta1.multivar.model)/Beta1.multivar.model*100
    delta.beta.list<-append(delta.beta.list,delta.Beta.perc)
    print(delta.Beta.perc)
  }
}
```

```{r}
df.results<-data.frame(row.names =comb.list)
df.results$Beta.perc.value<-delta.beta.list
df.results
```
#fbs/trestbps

```{r}
univar.model<-glm(as.formula(paste0('target~','fbs')),data=ds,family = binomial(link="logit"))
multi.model<-glm(as.formula(paste0('target~',paste0(c('fbs','trestbps'),collapse='+'))),data=ds,family = binomial(link="logit"))
multi.interact.model<-glm(as.formula(paste0('target~',paste0(c('fbs','trestbps'),collapse='+'),'+',paste0(c('fbs','trestbps'),collapse='*'))),data=ds,family = binomial(link="logit"))
summary(univar.model)
summary(multi.model)
summary(multi.interact.model)
```

```{r}
dich.variables<-c('sex')
contin.variables<-c('chol','thalach','oldpeak','trestbps')
covariates<-attr(glm.final_$coefficients,"names")
median.list<-as.numeric()
comb.list<-as.character()
delta.beta.list<-as.numeric()
for (d.var in dich.variables)
  {
  for (c.var in contin.variables)
    {
    model.name<-paste0('gm.50.',d.var,'.',c.var)
    print(model.name)
    ds[model.name]<-glm.final_$coefficients["(Intercept)"]
    for (cv in covariates)
      {
      if (cv!="(Intercept)")
      {
        if (cv!=d.var & cv!=c.var) 
        { 
          if (grepl( ':', cv, fixed = TRUE)==TRUE)
          {
           cv1<-strsplit(cv,split=':',fixed=TRUE)[[1]][1]
           cv2<-strsplit(cv,split=':',fixed=TRUE)[[1]][2]
           
           ds[model.name]<-ds[model.name]+glm.final_$coefficients[cv]*ds[cv1]*ds[cv2]
          }
          else
            {
            ds[model.name]<-ds[model.name]+glm.final_$coefficients[cv]*ds[cv]
            }
            
        }
      }
      

    }
    median.value<-median(ds[,model.name])
    curve.d1.c<-function(x){exp(median.value+glm.final_$coefficients[c.var]*x+glm.final_$coefficients[d.var]*1)/(1+exp(median.value+glm.final_$coefficients[c.var]*x+glm.final_$coefficients[d.var]*1))}
    curve.d0.c<-function(x){exp(median.value+glm.final_$coefficients[c.var]*x)/(1+exp(median.value+glm.final_$coefficients[c.var]*x))}
    curve.odds<-function(x){curve.d1.c/curve.d0.c}
    
    print(ggplot() +
    ggtitle(model.name)+
    geom_function(fun=curve.d1.c,col='red')+
    geom_function(fun=curve.d0.c)+
   
    ylim(0, 1)+
    xlim(0, 1.5*max(ds[c.var])))
    

  }

  }
```
```{r}
prob<-exp(glm.final_$coefficients['sex'])/(1+exp(glm.final_$coefficients['sex']))
prob

```

```{r}
plogis(glm.final_$coefficients['sex'])
```

```{r}
plogis(glm.final_$coefficients['sex'])/(1-plogis(glm.final_$coefficients['sex']))
```
```{r}
table(ds$exang,ds$target)
68*221/(443*267)
```

