---
title: "STA22"
output: html_notebook
---

```{r}
library(VGAM)
library(ggplot2)
library(MASS)
library(tidyverse)
library(caret)
library(nnet)
library(corrplot)
library(mlogit)
library(caret)
library(SciViews)
library(graphics)
library(ROCR)
library(ggplot2)
```
age : age
sex : sex
cp : chest pain
trestbps : resting blood pressure
chol : serum cholestoral [mg/dl]
fbs : fasting blood sugar
restecg : resting elctrocardiographic results
thalach: maximum heart rate achieved
exang : exercise induced angina
oldpeak : ST depression induced by exercice relatice to test
slope : slope of peak exercise
ca : number of major vessels colored by flouroscopy
thal : Thalassemia (less hemoglobin than normal)



```{r}
PATH_TO_FILE="C:\\Users\\a.lance\\Documents\\MASTER\\STA211\\STA212\\hearth_disease.csv"
```

```{r}
ds<-read.csv(PATH_TO_FILE,sep=";")
ds<-na.omit(ds)
```

```{r}
set.seed(1234)
sample <- sample(c(TRUE, FALSE), nrow(ds), replace=TRUE, prob=c(0.7,0.3))
ds.train  <- ds[sample, ]
ds.test   <- ds[!sample, ]
```



# Set univariate model for each parameters
```{r}
parameters<-c("age","sex","cp" ,"trestbps","chol","fbs","restecg","thalach","exang","oldpeak","slope" ,"ca","thal")
p_values = numeric()
p_values2 = numeric()
z_value<-numeric()
Beta<-numeric()
GList<-numeric()
wald.test.list<-numeric()
SE.list<-numeric()
LCI.list<-numeric()
UCI.list<-numeric()
for (param in parameters){
  
  fmod<-glm(reformulate(c(param), response='target'),data = ds.train,family = "binomial")
  nmod <- glm(target~1,data = ds.train, family = 'binomial')
  G= -2*ln(as.numeric(exp(logLik(nmod)))/as.numeric(exp(logLik(fmod))))
  GList<- append(GList,G)
  p_values2<-append(p_values2,pchisq(G, 1,lower.tail = FALSE))
  Beta1<-as.numeric(fmod$coefficients[2])
  Beta<-append(Beta,Beta1)
  SE<-as.numeric(sqrt(diag(vcov(fmod)))[2])
  SE.list<-append(SE.list,SE)
  z_value<-abs(Beta1/SE)
  wald_test_p_value<-2*pnorm(z_value,lower.tail = FALSE)
  wald.test.list<-append(wald.test.list,wald_test_p_value)
  LCI.list<-append( LCI.list,Beta1-1.96*SE)
  UCI.list<-append( UCI.list,Beta1+1.96*SE)
}

```

```{r}
df.results<-data.frame(row.names =parameters )
df.results$coeffs<-Beta
df.results$p_value<-p_values2
df.results$p_valueWT<-wald.test.list
df.results$SE<-SE.list
df.results$LCI<-LCI.list
df.results$UCI<-UCI.list
df.results
```

Le paramètres fbs est exclus

```{r}
parameters<-c("age","sex","cp" ,"trestbps","chol","restecg","thalach","exang","oldpeak","slope" ,"ca","thal")
glm.model<-glm(reformulate(parameters, response='target'),data = ds.train,family = "binomial")
coeff.list.2<-as.numeric(glm.model$coefficients)
summary(glm.model)
```


paramètres à exclure : age/restecg

```{r}
parameters<-c("sex","cp" ,"trestbps","chol","thalach","exang","oldpeak","slope" ,"ca","thal")
glm.model<-glm(reformulate(parameters, response='target'),data = ds.train,family = "binomial")
coeff.list.3<-as.numeric(glm.model$coefficients)
summary(glm.model)
```
```{r}
parameters_excluded<-c("fbs","age","restecg")

for (p in parameters_excluded){
  new_param<-append(parameters,p)
  glm.model2<-glm(reformulate(new_param, response='target'),data = ds.train,family = "binomial")
  print(summary(glm.model2))
}
```

# Vérification logits lineaire pour variables continues

```{r}
parametres_continus<-c('thalach','oldpeak','trestbps')

```

```{r}
glm.model$coefficients
```
```{r}
for ( p in parametres_continus){
lfit<-loess(ds.train$target~ds.train[[p]]) 
lgpred<-log(predict(lfit)/(1-predict(lfit))) 
plot(lgpred~ds.train[[p]])}
```
## Intéractions potentielles

```{r}
parameters<-c("sex","cp" ,"trestbps","chol","thalach","exang","oldpeak","slope" ,"ca","thal")

p.value.list<-as.numeric()
coeff.list<-as.numeric()
combination<-as.character()
for (i in 1:length(parameters))
#for (i in parameters)
  {
    for (j in i:length(parameters)) 
        {
        if (parameters[i]!=parameters[j])
          {
            factor1= paste(parameters[i],parameters[j],sep='*')
            factor2= paste0(parameters,collapse="+")
            factor<-paste0(factor2,"+",factor1)
            print(factor)
            glm.fit <- glm(as.formula(paste0("target~", factor)), data=ds.train,family = "binomial")
            summary_<-summary(glm.fit)
            coeff.list<-append(coeff.list,summary_$coeff[12,1])
            p.value.list<-append(p.value.list,summary_$coeff[12,4])
            combination<-append(combination,factor1)
          }
        }
  }
  

```


```{r}
df.results<-data.frame(row.names =combination )
df.results$coeffs<-coeff.list
df.results$p_value<-p.value.list
df.results<-subset(df.results,p_value<0.01)
df.results
```




## Sélection des intéractions

```{r}
significant_interactions<-rownames(df.results)
initial_parameters<-c("sex","cp" ,"trestbps","chol","thalach","exang","oldpeak" ,"ca","thal")
all_intercation_combinations<-do.call("c",lapply(seq_along(significant_interactions),function(i) utils::combn(significant_interactions,i,FUN=list)))
for (c in all_intercation_combinations)
  {
    formula_<-paste0("target~",paste0(initial_parameters,collapse="+"),"+",paste0(c,collapse = '+'))
    glm.final_ <- glm(as.formula(formula_), data=ds.train,family = "binomial")
    print(summary(glm.final_))
  }
```
## Modèle final

```{r}
formula_<-paste0("target~",paste0(initial_parameters,collapse="+"),"+",paste0(c("chol*oldpeak","oldpeak*ca","ca*thal"),collapse = '+'))
glm.final_ <- glm(as.formula(formula_), data=ds.train,family = binomial(link="logit"))
print(summary(glm.final_))
```


```{r}
confint(glm.final_)
```

## ROC CURVE

```{r}
pred<-predict(glm.final_,ds.train[parameters],type="response")
pred_ <- prediction(pred, ds.train$target)
perf <- performance(pred_,"tpr","fpr")
plot(perf,colorize=TRUE)
```

```{r}
auc_ROCR<-performance(pred_,measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
auc_ROCR
```
# Test performance

```{r}
pred<-predict(glm.final_,ds.test[parameters],type="response")
treshold<-0.5
Yfac   <- cut(ds.test$target, breaks=c(-Inf, treshold, Inf), labels=c("0", "1"))
yhatFac<-cut(pred,breaks = c(-Inf,treshold,Inf),lables=c('0','1'))
contingency.table<-table(Yfac,yhatFac)
addmargins(contingency.table)
sum(diag(contingency.table)) / sum(contingency.table)
```
## Interprétation
# ODD ratio

```{r}
dich.variables<-c('sex','fbs')
contin.variables<-c('chol','thalach','oldpeak','trestbps')
comb.list<-as.character()
delta.beta.list<-as.numeric()
for (d.var in dich.variables){
  for (c.var in contin.variables){
    comb.list<-append(comb.list,(paste0(c(d.var,c.var),collapse='*')))
    univar.model<-glm(as.formula(paste0('target~',d.var)),data=ds,family = binomial(link="logit"))
    multi.model<-glm(as.formula(paste0('target~',paste0(c(d.var,c.var),collapse='+'))),data=ds,family = binomial(link="logit"))
    multi.interact.model<-glm(as.formula(paste0('target~',paste0(c(d.var,c.var),collapse='+'),'+',paste0(c(d.var,c.var),collapse='*'))),data=ds,family = binomial(link="logit"))
    Beta1.univar.model<-as.numeric(univar.model$coefficients[d.var])
    Beta1.multivar.model<-as.numeric(multi.model$coefficients[d.var])
    delta.Beta.perc<-(Beta1.univar.model-Beta1.multivar.model)/Beta1.multivar.model*100
    delta.beta.list<-append(delta.beta.list,delta.Beta.perc)
    print(delta.Beta.perc)
  }
}
```

```{r}
df.results<-data.frame(row.names =comb.list)
df.results$Beta.perc.value<-delta.beta.list
df.results
```
#fbs/trestbps

```{r}
univar.model<-glm(as.formula(paste0('target~','fbs')),data=ds,family = binomial(link="logit"))
multi.model<-glm(as.formula(paste0('target~',paste0(c('fbs','trestbps'),collapse='+'))),data=ds,family = binomial(link="logit"))
multi.interact.model<-glm(as.formula(paste0('target~',paste0(c('fbs','trestbps'),collapse='+'),'+',paste0(c('fbs','trestbps'),collapse='*'))),data=ds,family = binomial(link="logit"))
summary(univar.model)
summary(multi.model)
summary(multi.interact.model)
```

```{r}
# target~sex+thalach
ds$gm50_sex_thalach<-glm.final_$coefficients['cp']*ds$cp+
                  glm.final_$coefficients['trestbps']*ds$trestbps+
                  glm.final_$coefficients['chol']*ds$chol+
                  #glm.final_$coefficients['thalach']*ds.train$thalach+
                  glm.final_$coefficients['exang']*ds$exang+
                  glm.final_$coefficients['oldpeak']*ds$oldpeak+
                  glm.final_$coefficients['ca']*ds$ca+
                  glm.final_$coefficients['thal']*ds$thal+
                  glm.final_$coefficients['chol:oldpeak']*ds$chol*ds$oldpeak+
                  glm.final_$coefficients['oldpeak:ca']*ds$ca*ds$oldpeak+
                  glm.final_$coefficients['ca:thal']*ds$ca*ds$thal+
                  glm.final_$coefficients['(Intercept)']
                  
  
  
gm50_median<-median(ds$gm50_sex_thalach)
thalach.median<-median(ds$thalach)

gm_sex1_chol<-function(x){exp(gm50_median+glm.final_$coefficients['thalach']*x+glm.final_$coefficients['sex'])/(1+exp(gm50_median+glm.final_$coefficients['thalach']*x+glm.final_$coefficients['sex']))}

gm_sex0_chol<-function(x){exp(gm50_median+glm.final_$coefficients['thalach']*x)/(1+exp(gm50_median+glm.final_$coefficients['thalach']*x))}



curve(gm_sex1_chol, col = 'red',from=0, to=300)
curve(gm_sex0_chol, col = 'green',from=0, to=300, add = TRUE)

```





